---
title: 'Building a Successful  KickStarter Campaign'
author: "Rachel Nelson"
date: "8/1/2020"
output:
  word_document: default
  pdf_document: default
---


#Section 1 
###Introduction: 
How can I make my KickStarter campaign a success? 

###Research questions
* Are there certain types/category of campaigns that are more successful?
* How much money should you ask for?
* Is there a time period for the campaign that works better than others?
* What is the average contribution of a backer?
* Is there a better time of year to launch a campaign?

###Approach
* I will be performing basic data analysis and correlation on the data set provided. I will  review things like the mean, median and mode of some of the factors that are of interest. 

####How your approach addresses (fully or partially) the problem.
By finding out which metrics matter, we can use these elements to ensure your next kickstarter campaign ends in success.

####Data 
https://www.kaggle.com/kemical/kickstarter-projects

####Required Packages
* dplyr 
* ggplot2
* plotly
* lm.beta


####Plots and Table Needs
* Scatter plots
* data tables
* correlation tables
* box plots

####Questions for future steps
* Should I look into neural networks? 

# Section 2
### How to import and clean my data
I am importing the data by connecting the the CSV that was available for download on the Kaggle site. 
https://www.kaggle.com/kemical/kickstarter-projects

```{r}

# load the data
ks_df <- read.csv("D:/College/DSC520/dsc520/data/ks-projects-201801.csv")

```

I am cleaning the data set to prepare it for analysis. 
####Check for missing columns
```{r}

# Check for Missing Columns
names(ks_df)
ks_df$rowid <- paste(ks_df$ID, "-", ks_df$round)
length(unique(ks_df$rowid))
length(ks_df$rowid)

```
Here I confirmed that all rows have a unique ID. I also reviewed the data to ensure all the data I needed was contained within the data set. 

####Check variables names
```{r}
# checks variable names and replace with new name
library(dplyr)
ks_df <- rename(ks_df, usd_pledged = usd.pledged)

```

Here I renamed the variable usd.pledged to usd_pledged to align the naming conventions of all of my headers, since the rest of the headers uses underscores instead of periods for spaces. 

####Check missing observations
```{r}
# checks for missing values in observations
colMeans(is.na(ks_df))
# removes column from data set
ks_df = subset(ks_df, select = -c(usd_pledged) )

```
Here I am looking for missing values. There is a small amount of data in the usd_pledged with missing values. If I wanted to cleanse the data set, I could remove these values, but for now, I want to keep it in mind since there are zero missing values from usd_pledged_real, which is a column giving the same information, but the conversion to USD was done from the fixer.io api instead of done by kickstarter. Instead of removing the rows with the missing data, I am going to remove the column from the data set since it is a duplicate column. 

usd_pledged: conversion in US dollars of the pledged column (conversion done by kickstarter).
usd pledge real: conversion in US dollars of the pledged column (conversion from Fixer.io API).


####Check variable classification
```{r}
# checks attributes of data frame
str(ks_df)

```
Checking the variable classification is the step used to make sure the data is the right datatype for analysis. 


####Check duplicate rows
```{r}

# Checking if one row is identical to another
distinctdata <- distinct(ks_df)
nrow(ks_df)
nrow(distinctdata)

```
Checking for duplicate rows within the data. None were found. If duplicate rows are found, the duplicate should be extracted from the dataset. 

####Change dates from factors to date
```{r}
ks_df <- transform(ks_df, deadline = as.Date(deadline), launched = as.Date(launched), backers = as.numeric(backers))
```
Changes the data type of deadline and launched to date.

### What does the final data set look like? 
```{r}
head(ks_df)
```





### Questions for future steps
I need to figure out if and how the factor/category data needs to be changed to numerical data. I also had to change dates from factors to date data types.

### What information is not self-evident?
I plan to run both correlation and unsupervised learning models on the data to see if I can uncover any new information that is not self-evident.


### What are different ways you could look at this data?
Yes, the questions I want to answer can be viewed though looking at bar charts, frequency plots and statistical models. 
* Are there certain types/category of campaigns that are more successful?
* How much money should you ask for?
* Is there a time period for the campaign that works better than others?
* What is the average contribution of a backer?
* Is there a better time of year to launch a campaign?

### How do you plan to slice and dice the data?.
Created a new variable for % successful by taking the pledged and dividing it by the goal. I also slided out the month for both deadline and launch dates. 
```{r}
# Adding new rows to slide and dice the data later
ks_df <-
  ks_df %>%
  mutate(
    pledged_to_goal = usd_pledged_real/usd_goal_real,
    count = 1,
    deadline_month = format(deadline,"%m"),
    launched_month = format(launched,"%m"),
    backers_per_pledge = usd_pledged_real/backers
    )

```


### How could you summarize your data to answer key questions?
This ties into the different ways I can look at the data set. Charts and visualizations are a great way to summarize the data and answer key questions. 

### What types of plots and tables will help you to illustrate the findings to your questions? 
Bar charts, box plots and scatter charts will help illustrate findings to my questions. 

### Do you plan on incorporating any machine learning techniques to answer your research questions? Explain.
Yes, I plan to see if there are any supervised (like decision tree or random forest) models and unsupervised (clustering) that can help make sense of what is funded verses unfunded. 

### Questions for future steps
This still ties in to question #3, where I need to figure out if the factor/categyory data needs to be changed to numerical data and if so, how I go about doing that. 

# Section 3
### Introduction
Kickstarter campaigns is a way to crowdsource funding to support projects, people or situations. It's a way to raise money. In this analysis, I will be finding out if there are controllable factors which can lead to a successful campaign. 

### The problem statement you addressed
Is there a way to design a kickstarter campaign to increase it's likelihood to be successful?

### How you addressed this problem statement 
I addressed this problem statement by looking into answering 5 questions: 
* Are there certain types/category of campaigns that are more successful?
* How much money should you ask for?
* Is there a time period for the campaign that works better than others?
* What is the average contribution of a backer?
* Is there a better time of year to launch a campaign?

I also performed correlation and applied machine learning techniques to see if there are ways to increase the likelihood of building successful campaigns. 

### Analysis

#### Are there certain types/category of campaigns that are more successful?
```{r}

# Filtering by one criterion
ks_dff <- filter(ks_df, state == "successful")

## Successful Launches based on Deadline Month
counts <- table(ks_dff$main_category)
barplot(counts, main="# of Successful Launches",
   xlab="Main Category")


```
The top 5 categories with successful campaigns are:
1. Music
2. Film & Video
3. Games
4. Publishing
5. Art



#### How much money should you ask for?
```{r}
library(ggplot2)
## Create a scatterplot of all states
ggplot(ks_df, aes(x=state, y=goal, col=state)) + ggtitle("State vs Goal Amount") + xlab("State") + ylab("Goal Amount") + geom_point(aes(colour = state))

# Boxplot of only successful campaigns
boxplot(goal~count,data=ks_dff, main="Successful Campaigns",
   xlab="Goal", ylab="Campaign")

summary(ks_dff)


```
Successful campaigns have a smaller range then non-successful campaigns. 
The average successful campaign has a goal of around 10,000 with a median of around 4,000.

#### Is there a time period for the campaign that works better than others?
```{r}
## Successful Launches based on Deadline Month
counts <- table(ks_dff$deadline_month)
barplot(counts, main="# of Successful Launches",
   xlab="Deadline Month")

```
May has the highest number of campaigns that are successful.


#### What is the average contribution of a backer?
```{r}

# Boxplot of only successful campaigns
boxplot(usd_pledged_real~count,data=ks_dff, main="Pledged Per Backer",
   xlab="Goal", ylab="Campaign")

summary(ks_dff)

```
The median backer pledges 63 USD to projects. 


#### Is there a better time of year to launch a campaign?
```{r}
## Successful Launches
counts <- table(ks_dff$launched_month)
barplot(counts, main="# of Successful Launches",
   xlab="Months")
```
March and October has the most for count of successful launches. December has the least. 

#### What are the factors that contribute to sucessful campaigns?
```{r}
# Prepping the data for modelling:

# Adding new rows to indicate successful campaigns
ks_dff <-
  ks_dff %>%
  mutate(
    successful = 1
    )

# Filtering by one criterion where campaigns not successful
ks_dfn <- filter(ks_df, state != "successful")


# Adding new rows to indicate unsuccessful campaigns
ks_dfn <-
  ks_dfn %>%
  mutate(
    successful = 0
    )

#combines successful and unsuccessful campaigns
df_union1<-merge(ks_dff,ks_dfn,all=TRUE)

df_union1 <- transform(df_union1, deadline_month = as.integer(deadline_month), launched_month = as.integer(launched_month))

model_1 <- lm(successful ~ backers+usd_pledged_real,usd_goal_real,pledged_to_goal+deadline_month+launched_month, data = df_union1)
summary(model_1)

library(lm.beta)
model_1.beta <- lm.beta(model_1)
coef(model_1.beta)

# linear regression on backers
linearMod <- lm(successful ~ backers, data=df_union1)
print(linearMod)
summary(linearMod)




```
The number of backers is a significant factor when predicting if the kickstarter will be a success.


### Conclusion
The best way to have a successful campaign is to increase the number of backers for that campaign. Would not recommend campaigning during the holiday season. 

