---
title: "7.1 Assignment"
author: "Rachel Nelson"
date: "7/18/2020"
output: 
  word_document: default
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
library(foreign)
library(caTools)
library(class)


```

a. What is the accuracy of the logistic regression classifier?


```{r}
# load the data
bcd_df <- read.csv("C:/Users/Rachel/Desktop/College/DSC520/dsc520/data/binary-classifier-data.csv")

#split data into test and train
split <- sample.split(bcd_df, SplitRatio = 0.8)
train.bcd <- subset(bcd_df, split == "TRUE")
test.bcd <- subset(bcd_df, split == "FALSE")


# build a model using all factors
newModel<-glm(label ~ x + y, data = train.bcd, family = binomial()) 
summary(newModel)

## run test through data
res <- predict(newModel, test.bcd, type = "response")
res <- predict(newModel, train.bcd, type = "response")

#validate the model
confmatrix <- table(Actual_value=train.bcd$label, Predicted_Value = res >0.5)
confmatrix

#Accuracy
(confmatrix[[1,1]]+confmatrix[[2,2]])/sum(confmatrix)

```
The accuracy of the logistic regression is 58%.

###b. How does the accuracy of the logistic regression classifier compare to the nearest neighbors algorithm?
```{r}
# load the data
bcd_df <- read.csv("C:/Users/Rachel/Desktop/College/DSC520/dsc520/data/binary-classifier-data.csv")
head(bcd_df)
str(bcd_df)

# creating test and training data sets 
bcd.subset <- as.data.frame(bcd_df)
head(bcd.subset)

nomalize <- function(x) {
  return ((x - min(x))/(max(x)-min(x)))}

bcd.subset.n <- as.data.frame(lapply(bcd.subset[,2:3], normalize)
head(bcd.subset.n)

set.seed(123)
dat.d <- sample(1:nrow(bcd.subset.n), size=nrow(bcd.subset.n)*0.7, replace = FALSE)


train.bcd <- bcd.subset[dat.d,]
test.bcd <- bcd.subset[-dat.d,]

train.bcd_labels <- bcd.subset[dat.d,1]
test.bcd_labels <- bcd.subset[-dat.d,1]

```


```{r}
NROW(train.bcd_labels)
## KNN model and accuracy

knn.38 <- knn(train=train.bcd, test=test.bcd, cl=train.bcd_labels, k=38)
ACC.38 <-100 * sum(test.bcd_labels == knn.38)/NROW(test.bcd_labels)
ACC.38
```
The accuracy of the KNN model is 97.3%

###c. Why is the accuracy of the logistic regression classifier different from that of the nearest neighbors
They are using different models/ways to predict the outcome. Good data scientists will always try different models with their data to find the one with the best fit. The KNN algorithm will predict the outcome based on the distance and average of plots near them. 

